Distributional Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain

### 2018-02-20 pkg/disttree/inst/draft/draft.Rnw
revised parts of the application section
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2031 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: ca4ee8f26f125d9b558b9052119f254f506eea3c

\Abstract{
While many regression models consider only the mean of the response variable recent advances in parametric modeling have made it possible to model all distribution parameters. In that way the underlying distribution can be fully specified allowing for a wide range of analysis and prediction methods which are particularly valuable in applications such as probabilistic weather forecasting. In generalized additive models for location, scale, and shape %(GAMLSS, \citealt{Rigby+Stasinopoulos:2005}) 
each distribution parameter can be modeled separately by its own generalized additive model (GAM) capturing linear and non-linear effects. A different approach that would make it possible to also model non-additive effects and interactions is to apply tree-structured models such as regression trees %(\citealt{Loh:2011}) 
or random forests. %(\citealt{Breiman:2001}). 
%%%<TH> "random forest" generisch; Breiman and Cutler's randomForest </TH>
This class of tree and forest models will now be integrated in probabilistic forecasting providing a novel general distributional framework for these methods. In particular, we are going to introduce distributional trees and forests which embed the idea of flexible parametric modeling into the framework of regression trees and random forests. 
%%%<TH> break-up very long sentence </TH>
Applying these novel approaches to precipitation data in a mountainous region with a very large number of numerical weather prediction quantities shows one of their main advantages in selecting variables automatically and illustrates their strengths as their performance is compared to GAMLSS models that are specified based either on subject-matter knowledge and experience from meteorology or based on a computationally more demanding boosting approach.
}
\Keywords{parametric models, regression trees, random forests, recursive partitioning, probabilistic forecasting, distributional regression}

@@

As explained in the introduction, distributional forests provide a new approach in probabilistic forecasting by embedding the ideas of flexible parametric modeling into the framework of tree-structured models. The ability to deal with a wide variety of effects and an automatic variable selection and detection of interactions are features provided by distributional forests that could be particularly valuable in probabilistic weather forecasting in order to tackle the above mentioned challenges.
To show their strengths in practice, distributional forests are now applied to probabilistic precipitation forecasts in a mountainous region based on a large number of numerical weather prediction quantities. Their results are compared to those of other zero-censored Gaussian models including a standard ensemble model output approach and two GAMLSS models. For one of the GAMLSS models the regressors have been selected and interactions have been specified in advance based on subject-matter knowledge and experience from meteorology. The other one is based on a computationally more demanding boosting approach.

@@

In particular, 14 basic forecast variables are considered with up to 12 variations of each of these variables being included as covariates. Since an ensemble of individual predictions is provided for each forecast variable, the possible variations include the mean, the standard deviation, the minimum and the maximum over all ensemble members, noted as ensemble mean, ensemble std. deviation, ensemble minimum and ensemble maximum. All of these values can be calculated on the sums over 24 hours (6--6 UTC of the following day) or a 6 hour time window. 
Another possible option is to consider the mean, minimum or maximum over 24 hours for each individual prediction and then take the ensemble mean or ensemble standard deviation, noted as ensemble mean of mean/minimal/maximal values or ensemble std. deviation of mean/minimal/maximal values respectively. Table~\ref{covariates} shows all variables together with the number of variations and the provided variations listed in the last column. Altogether 82~covariates are available in this setting.

@@

%\newpage
\begin{table}[t!]
%\resizebox{1.4\textwidth}{!}{
\begin{minipage}{\textwidth}
\begin{tabular}{ l  c  l }
\hline
Basic covariates & {\#} & Variations\\
\hline
\emph{tp}: total precipitation            & 12  & ensemble mean of sums over 24h, \\
\hspace*{0.5cm} power transformed (by $\frac{1}{1.6}$) &    & ensemble std. deviation of sums over 24h, \\ 
\emph{cape}: convective available     &    & ensemble minimum of sums over 24h, \\
\hspace*{0.9cm} potential energy &    & ensemble maximum of sums over 24h\\
\hspace*{0.9cm} power transformed (by $\frac{1}{1.6}$)&    & \qquad all for 6--30 UTC\\
\hspace*{\fill}                            &    & ensemble mean of sums over 6h\\
\hspace*{\fill}                            &    & \qquad for 6--12, 12--18, 18--24, 24--30 UTC\\
\hspace*{\fill}                            &    & ensemble std. deviation of sums over 6h\\
\hspace*{\fill}                            &    & \qquad for 6--12, 12--18, 18--24, 24--30 UTC\\
\hline
\emph{dswrf}: downwards short wave      & 6 & ensemble mean of mean values, \\
\hspace*{1.13cm} radiation flux (``sunshine'') &   & ensemble mean of minimal values,\\
\emph{msl}: mean sea level pressure     &   &  ensemble mean of maximal values,\\
\emph{pwat}: preciptable water          &   & ensemble std. deviation of mean values,\\
\emph{tmax}: 2m maximum temperature     &   & ensemble std. deviation of minimal values,\\
\hspace*{1.15cm}                          &   & ensemble std. deviation of maximal values,\\
\emph{tcolc}: total column-integrated   &   & \qquad all over 6--30 UTC \\
\hspace*{0.95cm} condensate               &   & \\
\emph{t500}: temperature on 500 hPa     &   & \\
\hspace*{1cm}                             &   & \\
\emph{t700}: temperature on 700 hPa     &   & \\
\hspace*{1cm}                             &   & \\
\emph{t850}: temperature on 850 hPa     &   & \\
\hspace{1cm}                              &   & \\
\hline
\emph{tdiff500850}: temperature         & 3 & ensemble mean of difference in mean,\\
\hspace*{\fill} difference 500 to 850 hPa &   & ensemble minimum of difference in mean,\\
\emph{tdiff500700}: temperature         &   & ensemble maximum of difference in mean\\
\hspace*{\fill} difference 500 to 700 hPa &   & \qquad all over 6--30 UTC\\
\emph{tdiff700850}: temperature         &   & \\
\hspace*{\fill} difference 700 to 850 hPa &   & \\
\hline
\emph{msl{\_}diff}: mean sea level pressure & 1 & \emph{msl{\_}mean{\_}max} $-$ \emph{msl{\_}mean{\_}min}\\
\hspace{1.6cm} difference            &   & \qquad over 6--30 UTC\\
\hline
\end{tabular}
\caption[Table caption text]{Basic covariates together with the number ({\#}) and the type of variations.}
\label{covariates}
\end{minipage} 
%}
\end{table}
\fixme{dswrf\_mean/sprd\_min not included in formula but in the 82 covariates}

@@

\item A GAMLSS where for each distribution parameter the selection of covariates and the specification of interactions is done in advance, both based on meteorological expert knowledge following \cite{Stauffer+Umlauf+Messner:2017} (prespecified GAM). Among the available 82 regressors, 9 where selected for the location parameter and 6 for the scale parameter. Additionally, for each distribution parameter one interaction term is specified in which convective available energy is interacting with total precipitation (with the ensemble mean for the location parameter and with the ensemble standard deviation for the scale parameter). 
%In this terms an ensemble mean of 6 hour sums of convective available energy is interacting with an ensemble mean or standarad deviation of 6 hour sums of total precipitation.  
These interactions might capture events with a high potential for moderate or severe thunderstorms with large amounts of precipitation. By putting extra effort in the specification of the formulas the prespecified GAM model is provided with additional information (and is therefore expected to profit from this advantage).
\item A boosted GAMLSS \citep{Hofner+Mayr+Schmid:2016} using non-homogeneous boosting
\citep{Messner+Mayr+Zeileis:2017} for an automatic selection of covariates (boosted GAM). The idea of this (boosting) approach is to iteratively minimize a loss function where in each step only the coefficient leading to the highest improvement of the current model is updated. As the initial coefficients are all zero, only the most important covariates are included with nonzero coefficients while less relevant variables are ignored if the procedure is stopped before convergence.
\end{itemize}

@@

\subsection{Application for one station}
\label{Application for one station}
First of all we focus on one observation station, Axams in Tyrol. This station was selected due to its geographical closeness to Innsbruck, the capital of Tyrol. As for all other stations, daily observations and numerical weather predictions are available for the month of July of 28 years from 1985 through the year 2012.
In Figure~\ref{axams_onestation} the results of fitting a distributional forest to the first 24 years of the data set and predicting the total amount of precipitation for one specific day of each of the four successive years have been shown as a motivational example for a probabilistic forecast. Staying with this fitted model we are now going to analyze variable importance of the distributional forest model and compare its results to those of the GAM models and the EMOS model based on residual QQ-plots.
%%% <TH> wouldn't be a uniform QQ-plot be more appropriate? </TH>

%%%<TH> always us the term "distributional forest"
As shown in Table~\ref{model_specification} the distributional forest includes all 82~covariates. A single tree model would allow to easily analyze which covariates are used as split variables in the nodes and hence have a significant impact on the response variable. In a forest model though, for each node in each of the trees a subset of the covariates is chosen randomly as possible split variables to select one from. Moreover, each tree is built on a resampled version of the learning data. For these reasons the choices of split variables vary over the trees of a forest making it difficult to analyze the impact of covariates just by investigating the different trees. To get an overview of which covariates have the most influential effect on the response variable a standard measurement of variable importance for forest models has been applied. Based on the mean decrease in accuracy regarding the CRPS the 10 leading covariates of the above described distributional forest for station Axams learned on the first 24 years are listed in Figure~\ref{varimp}. As expected, the numerical forecast for total precipitation is the most influential variable. In particular, five variations of this variable are among the top ten of the list of covariates ordered by variable importance.

@@

In Figure~\ref{qqr_oosample} another impression of the quality of the predictions made by the tested methods is given by residual QQ-plots for out-of-sample predictions of each of the models. For these plots 100 randomized quantile residuals are simulated per observation. The closer the points are to the diagonal line the better the estimated distributional model fits the data. Overall, it can be stated that all four models provide well calibrated forecasts. This is also supported by the corresponding probability integral transform (PIT) histograms (\citealt{Gneiting+Balabdaoui+Raftery:2007}) in the appendix.

@@

After analyzing the results of the models on only one data set where visualization is easier, we now want to get a more profound impression of the performances of the different methods. For that reason the four methods are evaluated in a cross validation framework for the station Axams. The data set over all 28 years is split into 7 subsets each consisting of 4 randomly selected years. While 6 of these subsets are used as learning data the 7th is the testing data set for which predictions are made and evaluated. This is done by calculating the CRPS value for each of the observations in the testing data set and then averaging them. Each of the 7 parts is selected as testing data once which leads to 7 different settings of which again the average values are stored. This whole procedure is repeated 10 times. The resulting skill scores are illustrated in Figure~\ref{boxplots_crps_axams} using the EMOS model as the reference model represented by the horizontal line at height 0. This CRPS skill score boxplot shows that, in this setting, both GAM models and the distributional forest perform better than the EMOS model. While the two GAM models lead to an improvement of around 4 percent with the boosted version being slightly ahead, the distributional forest even reaches an improvement rate close to 6 percent.
<<echo=FALSE, results=hide>>=
#### cross validation rain
if(file.exists("rain_Axams_7x10.rda")){
  load("rain_Axams_7x10.rda")
} else {
  source("rain_cross_7x10.R")
  library("gamlss.cens")
  gen.cens("NO", type = "left")
  rainres <- rain_cross(stationname = "Axams", 
                        seedconst = 7, ntree = 100,
                        forest_mtry = 27,
                        gamboost_cvr = TRUE,
                        frac = FALSE)

@@

\SweaveOpts{eval = TRUE}
%\newpage
\subsection{Application for all stations}
Until now the focus has been on one observation station only. To test and compare the four methods under different (geographical) circumstances we are now going to apply them on all 95 stations. At each of them, the first 24 years are used as learning data set and predictions are made for the ensuing 4 years of available observations. For these out-of-sample predictions the CRPS skill score with the EMOS model as reference model is calculated for each method at each station. In Figure~\ref{boxplot_crps_all} the boxplots of the CRPS skill score values are plotted and show that the distributional forest leads to the highest improvement averaged over all stations. The green line represents the results for the station Axams.
%NAs in gamlss in 13 stations!!!\\
%for example in 13th station in complete{\_}stations\\
%\text{Error in glim.fit(f = mu.object, X = mu.X, y = y, w = w, fv = mu, os = mu.offset,  :}\\ 
%\text{NA's in the working vector or weights for parameter mu}
%\\
% sum(is.na(crps[,3]))
%<<eval = TRUE, echo = FALSE>>=
<<echo = FALSE>>=
#### prediction over all stations 24 - 4
if(file.exists("rain_pred_24to4.rda")){
  #load("~/svn/partykit/pkg/disttree/inst/draft/rain_pred_24to4.rda")
  load("rain_pred_24to4.rda")
} else {

@@

Figure~\ref{map} shows a map of Tyrol with the location of all observation stations included in the data set. Each station is illustrated by a symbol with the type of symbol representing the method that performed best in terms of CRPS at this station. The color of the symbol indicates the difference between the CRPS of the distributional forest and the CRPS of the best performing among the other three models (prespecified GAM, boosted GAM and EMOS).
Even though for some stations, especially in East Tyrol, one of the two GAMs and for a few others the EMOS model leads to the best results, for the majority of stations the distributional forest performes best.

@@

From these results we can conclude that a distributional forest provides a powerful alternative framework requiring only little effort to set up a model which is competitive to commonly used methods or even outperforms them. 
For the prespecified GAM model the covariates needed to be selected and interactions needed to be defined in advance. To do so, meteorological experience and knowledge is required and a lot of thought has to be put in these decisions. Moreover, making such decisions is always equivalent to influencing the resulting model and comes along with the risk of having missed important facts or even imposing certain effects on the model which might not correspond to the truth. This risk is avoided in the boosted GAM as all covariates are included and the selection of the influential covariates is done within a boosting framework. However, the trade off here is a remarkably higher computational effort. To build a distributional forest the user does not have to put any additional effort in it and therefore is not required to provide any expert knowledge or additional information about the setting. Variable selection is done automatically and interactions are detected by the tree structure of the model. In this particular setting, a very high number of covariates is available which intensifies the impact of the above mentioned strength of the distributional forest model. Therefore, in comparing these methods the strengths of distributional forests are clearly outlined as they offer a very flexible framework which enables an easy use of the method and at the same time lead to an improvement in the predictive performance.

>>>>

\Abstract{
While many regression models consider only the mean of the response variable recent advances in parametric modeling have made it possible to model all distribution parameters. In that way the underlying distribution can be fully specified allowing for a wide range of analysis and prediction methods which are particularly valuable in applications such as probabilistic weather forecasting. In generalized additive models for location, scale, and shape %(GAMLSS, \citealt{Rigby+Stasinopoulos:2005}) 
each distribution parameter can be modeled separately by its own generalized additive model (GAM) capturing linear and non-linear effects. A different approach that would make it possible to also model non-additive effects and interactions is to apply tree-structured models such as regression trees %(\citealt{Loh:2011}) 
or random forests. %(\citealt{Breiman:2001}). 
%%%<TH> "random forest" generisch; Breiman and Cutler's randomForest </TH>
This class of tree and forest models will now be integrated in probabilistic forecasting providing a novel general distributional framework for these methods. In particular, we are going to introduce distributional trees and forests which embed the idea of flexible parametric modeling into the framework of regression trees and random forests. 
%%%<TH> break-up very long sentence </TH>
Applying these novel approaches to precipitation data in a mountainous region with a very large number of numerical weather prediction quantities shows one of their main advantages in selecting variables automatically and illustrates their strengths as their performance is compared to GAMLSS that are specified based either on subject-matter knowledge and experience from meteorology or based on a computationally more demanding boosting approach.
}
\Keywords{parametric models, regression trees, random forests, recursive partitioning, probabilistic forecasting, distributional regression}

@@

As explained in the introduction, distributional forests provide a new approach in probabilistic forecasting by embedding the ideas of flexible parametric modeling into the framework of tree-structured models. The ability to deal with a wide variety of effects and an automatic variable selection and detection of interactions are features provided by distributional forests that could be particularly valuable in probabilistic weather forecasting in order to tackle the above mentioned challenges.
To show their strengths in practice, distributional forests are now applied to probabilistic precipitation forecasts in a mountainous region based on a large number of numerical weather prediction quantities. Their results are compared to those of other zero-censored Gaussian models including a standard ensemble model output approach and two GAMLSS. For one of the GAMLSS the regressors have been selected and interactions have been specified in advance based on subject-matter knowledge and experience from meteorology. The other one is based on a computationally more demanding boosting approach.

@@

In particular, 14 basic forecast variables are considered with up to 12 variations of each of these variables being included as covariates. Since an ensemble of individual predictions is provided for each forecast variable, the possible variations include the mean, the standard deviation, the minimum and the maximum over all ensemble members, noted as ensemble mean, ensemble std. deviation, ensemble minimum and ensemble maximum. All of these values can be calculated on the sums over 24 hours (6--6 UTC of the following day) or a 6 hour time window. 
Another possible option is to consider the mean, minimum or maximum over 24 hours for each individual prediction and then take the ensemble mean or ensemble standard deviation, noted as ensemble mean of mean/minimal/maximal values or ensemble std. deviation of mean/minimal/maximal values respectively. Table~\ref{covariates} shows all variables together with the number of variations and the provided variations listed in the last column. Altogether 82~covariates are available in this setting. \fixme{Mention that dswrf\_mean/sprd\_min is not included?}

@@

%\newpage
\begin{table}[t!]
%\resizebox{1.4\textwidth}{!}{
\begin{minipage}{\textwidth}
\begin{tabular}{ l  c  l }
\hline
Basic covariates & {\#} & Variations\\
\hline
\emph{tp}: total precipitation            & 12  & ensemble mean of sums over 24h, \\
\hspace*{0.5cm} power transformed (by $\frac{1}{1.6}$) &    & ensemble std. deviation of sums over 24h, \\ 
\emph{cape}: convective available     &    & ensemble minimum of sums over 24h, \\
\hspace*{0.9cm} potential energy &    & ensemble maximum of sums over 24h\\
\hspace*{0.9cm} power transformed (by $\frac{1}{1.6}$)&    & \qquad all for 6--30 UTC\\
\hspace*{\fill}                            &    & ensemble mean of sums over 6h\\
\hspace*{\fill}                            &    & \qquad for 6--12, 12--18, 18--24, 24--30 UTC\\
\hspace*{\fill}                            &    & ensemble std. deviation of sums over 6h\\
\hspace*{\fill}                            &    & \qquad for 6--12, 12--18, 18--24, 24--30 UTC\\
\hline
\emph{dswrf}: downwards short wave      & 6 & ensemble mean of mean values, \\
\hspace*{1.13cm} radiation flux (``sunshine'') &   & ensemble mean of minimal values,\\
\emph{msl}: mean sea level pressure     &   &  ensemble mean of maximal values,\\
\emph{pwat}: preciptable water          &   & ensemble std. deviation of mean values,\\
\emph{tmax}: 2m maximum temperature     &   & ensemble std. deviation of minimal values,\\
\hspace*{1.15cm}                          &   & ensemble std. deviation of maximal values,\\
\emph{tcolc}: total column-integrated   &   & \qquad all over 6--30 UTC \\
\hspace*{0.95cm} condensate               &   & \\
\emph{t500}: temperature on 500 hPa     &   & \\
\hspace*{1cm}                             &   & \\
\emph{t700}: temperature on 700 hPa     &   & \\
\hspace*{1cm}                             &   & \\
\emph{t850}: temperature on 850 hPa     &   & \\
\hspace{1cm}                              &   & \\
\hline
\emph{tdiff500850}: temperature         & 3 & ensemble mean of difference in mean,\\
\hspace*{\fill} difference 500 to 850 hPa &   & ensemble minimum of difference in mean,\\
\emph{tdiff500700}: temperature         &   & ensemble maximum of difference in mean\\
\hspace*{\fill} difference 500 to 700 hPa &   & \qquad all over 6--30 UTC\\
\emph{tdiff700850}: temperature         &   & \\
\hspace*{\fill} difference 700 to 850 hPa &   & \\
\hline
\emph{msl{\_}diff}: mean sea level pressure & 1 & \emph{msl{\_}mean{\_}max} $-$ \emph{msl{\_}mean{\_}min}\\
\hspace{1.6cm} difference            &   & \qquad over 6--30 UTC\\
\hline
\end{tabular}
\caption[Table caption text]{Basic covariates together with the number ({\#}) and the type of variations.}
\label{covariates}
\end{minipage} 
%}
\end{table}

@@

\item A GAMLSS where for each distribution parameter the selection of covariates and the specification of interactions is done in advance, both based on meteorological expert knowledge following \cite{Stauffer+Umlauf+Messner:2017} (prespecified GAM). Among the available 82 regressors, 9 where selected for the location parameter and 6 for the scale parameter. Additionally, for each distribution parameter one interaction term is specified in which convective available energy is interacting with total precipitation (with the ensemble mean for the location parameter and with the ensemble standard deviation for the scale parameter). 
%In this terms an ensemble mean of 6 hour sums of convective available energy is interacting with an ensemble mean or standarad deviation of 6 hour sums of total precipitation.  
These interactions might capture events with a high potential for moderate or severe thunderstorms with large amounts of precipitation. By putting extra effort in the specification of the formulas the prespecified GAM is provided with additional information and is therefore expected to profit from this advantage.
\item A boosted GAMLSS \citep{Hofner+Mayr+Schmid:2016} using non-homogeneous boosting
\citep{Messner+Mayr+Zeileis:2017} for an automatic selection of covariates (boosted GAM). The idea of this boosting approach is to iteratively minimize a loss function where in each step only the coefficient leading to the highest improvement of the current model is updated. As the initial coefficients are all zero, only the most important covariates are included with nonzero coefficients while less relevant variables are ignored if the procedure is stopped before convergence.
\end{itemize}

@@

\subsection{Application for one station}
\label{Application for one station}
First of all we focus on one observation station, Axams in Tyrol. This station was selected due to its geographical closeness to Innsbruck, the capital of Tyrol. As for all other stations, daily observations and numerical weather predictions are available for the month of July of 28 years from 1985 through the year 2012.
In Figure~\ref{axams_onestation} the results of fitting a distributional forest to the first 24 years of the data set and predicting the total amount of precipitation for one specific day of each of the four successive years have been shown as a motivational example for a probabilistic forecast. Staying with this fitted model we are now going to analyze variable importance of the distributional forest model and compare its results to those of the GAMs and the EMOS model based on residual QQ-plots.
%%% <TH> wouldn't be a uniform QQ-plot be more appropriate? </TH>

%%%<TH> always us the term "distributional forest"
As shown in Table~\ref{model_specification} the distributional forest includes all of the available covariates. A single tree model would allow to easily analyze which covariates are used as split variables in the nodes and hence have a significant impact on the response variable. In a forest model though, for each node in each of the trees a subset of the covariates is chosen randomly as possible split variables to select one from. Moreover, each tree is built on a resampled version of the learning data. For these reasons the choices of split variables vary over the trees of a forest making it difficult to analyze the impact of covariates just by investigating the different trees. To get an overview of which covariates have the most influential effect on the response variable a standard measurement of variable importance for forest models has been applied. Based on the mean decrease in accuracy regarding the CRPS the 10 leading covariates of the above described distributional forest for station Axams learned on the first 24 years are listed in Figure~\ref{varimp}. As expected, the numerical forecast for total precipitation is the most influential variable. In particular, five variations of this variable are among the top ten of the list of covariates ordered by variable importance.

@@

In Figure~\ref{qqr_oosample} an impression of the quality of predictions made by the tested methods is given by residual QQ-plots for out-of-sample predictions of each of the models. For these plots 100 randomized quantile residuals are simulated per observation. The closer the points are to the diagonal line the better the estimated distributional model fits the data. Overall, it can be stated that all four models provide well calibrated forecasts. This is also supported by the corresponding probability integral transform (PIT) histograms (\citealt{Gneiting+Balabdaoui+Raftery:2007}) in the appendix.

@@

After analyzing the results of the models on only one data set where visualization is easier, we now want to get a more profound impression of the performances of the different methods. For that reason the four methods are evaluated in a cross validation framework for the station Axams. The data set over all 28 years is split into 7 subsets each consisting of 4 randomly selected years. While 6 of these subsets are used as learning data the 7th is the testing data set for which predictions are made and evaluated. This is done by calculating the CRPS value for each of the observations in the testing data set and then averaging them. Each of the 7 parts is selected once as testing data which leads to 7 different settings of which again the average values are stored. This whole procedure is repeated 10 times. The resulting skill scores are illustrated in Figure~\ref{boxplots_crps_axams} using the EMOS model as the reference model represented by the horizontal line at height 0. This CRPS skill score boxplot shows that, in this setting, both GAMs and the distributional forest perform better than the EMOS model. While the two GAMs lead to an improvement of around 4 percent with the boosted version being slightly ahead, the distributional forest even reaches an improvement rate close to 6 percent.
<<echo=FALSE, results=hide>>=
#### cross validation rain
if(file.exists("rain_Axams_7x10.rda")){
  load("rain_Axams_7x10.rda")
} else {
  source("rain_cross_7x10.R")
  library("gamlss.cens")
  gen.cens("NO", type = "left")
  rainres <- rain_cross(stationname = "Axams", 
                        seedconst = 7, ntree = 100,
                        forest_mtry = 27,
                        gamboost_cvr = TRUE,
                        frac = FALSE)

@@

\SweaveOpts{eval = TRUE}
%\newpage
\subsection{Application for all stations}
Until now the focus has been on one observation station only. To test and compare the four methods under different geographical circumstances we are now going to apply them on all 95 stations. At each of them, the first 24 years are used as learning data set and predictions are made for the ensuing 4 years of available observations. For these out-of-sample predictions the CRPS skill score with the EMOS model as reference model is calculated for each method at each station. In Figure~\ref{boxplot_crps_all} the boxplots of the CRPS skill score values are plotted and show that the distributional forest leads to the highest improvement averaged over all stations. The green line represents the results for the station Axams.
%NAs in gamlss in 13 stations!!!\\
%for example in 13th station in complete{\_}stations\\
%\text{Error in glim.fit(f = mu.object, X = mu.X, y = y, w = w, fv = mu, os = mu.offset,  :}\\ 
%\text{NA's in the working vector or weights for parameter mu}
%\\
% sum(is.na(crps[,3]))
%<<eval = TRUE, echo = FALSE>>=
<<echo = FALSE>>=
#### prediction over all stations 24 - 4
if(file.exists("rain_pred_24to4.rda")){
  #load("~/svn/partykit/pkg/disttree/inst/draft/rain_pred_24to4.rda")
  load("rain_pred_24to4.rda")
} else {

@@

Figure~\ref{map} shows a map of Tyrol with the location of all observation stations included in the data set. Each station is illustrated by a symbol with the type of symbol representing the method that performed best in terms of CRPS at this station. The color of the symbol indicates the difference between the CRPS of the distributional forest and the CRPS of the best performing among the other three models (prespecified GAM, boosted GAM and EMOS).
Even though for some stations, especially in East Tyrol, one of the two GAMs and for a few others the EMOS model leads to the best results, for the majority of stations the distributional forest performs best.

@@

From these results we can conclude that a distributional forest provides a powerful alternative framework requiring only little effort to set up a model which is competitive to commonly used methods or even outperforms them. 
Even though a proper selection of covariates and specification of interactions in advance can lead to better fitting (and probably less complex) models, this presupposes profound knowledge of all influencing factors. Otherwise, important aspects might be missed or effects which do not correspond to the truth might be imposed on the model. This risk can be avoided by applying a boosting algorithm for an automatic variable selection. In this case the trade-off is often a remarkably higher computational effort. 
To build a distributional forest the user is not required to provide any expert knowledge or additional information about the setting as variable selection and detection of interactions are done automatically by the tree-structure of the model. This feature is particularly advantageous in case of a very high number of covariates. Therefore, the above discussed application illustrates the strengths of distributional forests as they offer a very flexible framework which enables an easy use of the method and at the same time lead to an overall improvement in the predictive performance in this setting.
